---
title: "Using TEMPO"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using TEMPO}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

`TEMPO` provides access to the Romanian National Institute of Statistics'
TEMPO Online database. It has two public functions:

* `tempo_toc()` lists the available statistical tables and their matrix codes.
* `tempo_bulk()` downloads one or more complete tables as CSV files.

The database is an external service, so the examples below need an internet
connection. They are deliberately not evaluated when this vignette is built.

## Find a table

Start by retrieving the table of contents. The default language is Romanian;
set `language = "en"` to obtain English table names.

```{r list-tables, eval = FALSE}
library(TEMPO)

tables <- tempo_toc(language = "en")
head(tables)
```

The result is a data frame with `name` and `code` columns. Search the `name`
column to identify a table of interest, then retain its code.

```{r search-tables, eval = FALSE}
population_tables <- tables[
  grepl("population", tables$name, ignore.case = TRUE),
  c("name", "code")
]
head(population_tables)
```

Use `full_description = TRUE` when you need the statistical domain,
sub-domain, survey name, and last-update date for every listed table. This
makes one additional request for every matrix, so it can take a long time.

```{r detailed-toc, eval = FALSE}
table_details <- tempo_toc(full_description = TRUE, language = "en")
head(table_details)
```

## Download a table

Choose an output directory that you control. `tempo_bulk()` creates it if
needed and writes only there. For example, the code below downloads the
`ACC101B` matrix in English.

```{r download-one, eval = FALSE}
data_directory <- file.path(tempdir(), "tempo-data")

files <- tempo_bulk(
  codes = "ACC101B",
  language = "en",
  directory = data_directory
)

files
```

`tempo_bulk()` returns the paths of files it downloaded, invisibly. A repeated
call returns no path when the local file is already current, so construct the
CSV path from the output directory and matrix code when later steps need the
file. The downloaded files are ordinary comma-separated CSV files.

```{r read-csv, eval = FALSE}
csv_path <- file.path(data_directory, "ACC101B.csv")
accidents <- utils::read.csv(
  csv_path,
  stringsAsFactors = FALSE,
  check.names = FALSE
)
utils::head(accidents)
```

## Download several tables

Pass a character vector to download multiple matrices. Requests are made
sequentially, which keeps use of the public service conservative.

```{r download-many, eval = FALSE}
codes <- c("ACC101B", "ACC101C")
files <- tempo_bulk(codes, language = "ro", directory = data_directory)
```

Some matrices have dimensions with many selectable values. `TEMPO` splits
those requests into sequential chunks and combines the resulting CSV content
into one file per matrix. A complete table can still be large, so download
only the tables you need.

## Reuse local files

Before replacing a requested file, `tempo_bulk()` compares its modification
date with the update date published by TEMPO Online. Repeating a download is
therefore safe: files that are already current are skipped.

```{r incremental, eval = FALSE}
tempo_bulk("ACC101B", language = "en", directory = data_directory)
#> Skipping `ACC101B`: the local CSV is up to date.
```

## Handle a temporarily unavailable service

The service can be unavailable or its response format can change. In that
case, the functions stop with an informative error. For longer scripts, wrap
the call in `tryCatch()` and decide how your workflow should recover.

```{r error-handling, eval = FALSE}
tables <- tryCatch(
  tempo_toc(language = "en"),
  error = function(error) {
    message("TEMPO Online is unavailable: ", conditionMessage(error))
    NULL
  }
)
```
